Imprecise Datapath Design: An Overclocking Approach

### 2013-08-05 TRETS2013/TRETS_overclocking.tex
Minor changes

\keywords{Overclocking, Imprecise Design, Approximate Computing}

@@

\section{Background}\label{Section_Background}
\subsection{Word-length Optimization}\label{Section_Background_WL}
There exists a significant amount of work demonstrating that optimizing the datapath precision would bring a substantial benefits on clock speed, silicon area and power consumption. There are two main research directions in this area. The first discusses methods to analyze the errors caused by the usage of finite precision and compute bounds on the worst-case errors seen at the output of the design. Related techniques include the simulation-based approach~\cite{SungWLop_Simulation} and the analytical-based approach such as interval arithmetic~\cite{moore1966IA}, affine arithmetic~\cite{de2004affine}, Satisfiability-Modulo Theories~\cite{SAT2010} and polynomial representations~\cite{Boland2011TCAD}. A detailed summary of these methods can be found in~\cite{GeorgeDT11}. In the second research field, these tools are utilized to choose the minimum precision within a datapath to satisfy a required accuracy specification~\cite{RoldaoWLop_Simulation}. In this work, two different design scenarios are analyzed. In the conventional scenario, we speculate and reduce the original word-length in the datapath under a given timing constraint while ensuring no timing errors will occur. In the proposed new scenario, we operate the datapath with the original word-length while allowing timing violations sometimes to occur. The truncation errors derived in the former scenario are then compared to the overclocking errors generated in our proposed new scenario.

@@

\subsection{Imprecise Circuit Design Methodology}\label{Section_Background_Imprecise}
It is worth noting that the choice of precision is not the only source of error when designing a datapath. Recently we have seen a growth of parallel streams of research which aim at exploring alternative methods to trade accuracy for design efficiency. This strand of research is motivated by the fact that extra benefits of manufacturing, test, power and timing can be obtained if the tight requirement of absolute correctness is released for devices and interconnect, as pointed out by the International Technology Roadmap for Semiconductors (ITRS) in 2007~\cite{semiconductor2007international}. This topic is expected to be of growing importance in the future, because we will face new design challenges as the technology scales further.

@@

\subsection{Determination of the Optimum Adder Structure for Given Frequency Requirements}
First, suppose the algorithm designer wished to create a circuit that can run at a given frequency in a given resource constraint while achieving the minimum possible output error. A decision must then be taken over which adder structure achieves this minimum, and whether the structure should be overclocked

@@

We first notice that for all frequency values, the overclocked RCA achieves no larger error expectation than the RCA with truncated operand word-lengths, as predicted by the models for the RCA in Section~\ref{section_RCA_Comparison}. It can also be observed that the CSA cannot be implemented with the original word-length due to the limited area budget and this leads to large truncation errors. Although the CSA with 4 stages is best for some frequencies, the overclocked RCA is still the optimum design when high operating frequencies are applied.

We then perform similar experiments with a variety of area constraints. The optimum design methods with respect to different operating frequencies and area consumptions are demonstrated in Fig.~\ref{Fig_CSA_Tradeoff}. From this figure, several observations can be made. Firstly, if the frequency requirement is not large whilst the available area is large enough to implement a CSA in full precision, it will be the optimum design. This is expected from our earlier analysis in Fig.~\ref{Fig_CSA2stage_Timing}. Secondly, the 2-stage CSA is better than the 3-stage CSA and the 4-stage CSA when frequency is low, as it consumes less area although all achieve the same error expectation. Thirdly, only part of the CSA can be implemented under a tighter area budget, whereas the RCA still keeps full precision. In this case, area becomes the dominant factor and precision is lost for the CSA, meaning the RCA with overclocking is the optimum design method across almost the whole frequency domain in this situation. Finally, for very stringent area constraints, the word-length of RCA is also limited. This results in truncation errors initially for all design scenarios. However, the RCA with overclocking can still be employed as the optimum design, because it loses less precision than the CSA under the same area constraint.
%
\begin{figure}[t]
  \centering
  \vspace{-2.5ex}
  \includegraphics[width=4.5in]{./Figures/Tradeoff_Freq.eps}
  \vspace{-2.5ex}
  \caption{A demonstration of the optimum design methodology which achieves the minimum error at outputs with respect to a variety of frequency and area constraints.}
  \label{Fig_CSA_Tradeoff}
\end{figure}

\subsection{Determination of the Optimum Adder Structure for Given Accuracy Requirements}
If the design goal is to operate the circuit as fast as possible with the minimum area whilst a certain error budget can be tolerated, the optimum design methodology can be decided as illustrated in Fig.~\ref{Fig_CSA_Tradeoff_Error}. In this situation, the error specifications are evaluated in terms of the mean relative error (MRE), as given by~(\ref{Eq:MRE}), where $E_{error}$ and $E_{out}$ refer to the mean value of error and the mean value of outputs, respectively.
%
\begin{eqnarray}\label{Eq:MRE}
  MRE=\left|\frac{E_{error}}{E_{out}}\right|\times 100\%
\end{eqnarray}

In our experiments, MRE is set ranging from $0.001\%$ to $50\%$. For a certain value of MRE, the design with the maximum operating frequency is selected as the optimum design. Moreover, the smallest design is the optimum one if multiple structures operate at the same frequency, with a certain area requirement. Fig.~\ref{Fig_CSA_Tradeoff_Error} can thus be obtained based on these criteria.
%
\begin{figure}[t]
    \centering
    \vspace{-2ex}
    \includegraphics[width=4.5in]{./Figures/Tradeoff_Error2.eps}
    \vspace{-2.5ex}
    \caption{A demonstration of the optimum design methodology which runs at the fastest frequency with respect to a variety of accuracy and area constraints.}
    \label{Fig_CSA_Tradeoff_Error}
    \vspace{-1ex}
\end{figure}

For a tight accuracy requirement, i.e. $MRE<0.005\%$, CSA serves as the optimum design choice for large area constraints, as it intrinsically operates faster than RCA. Once again, when the area budget shrinks, the RCA performs best because the precision of the CSA is limited. Similarly to the results in Fig.~\ref{Fig_CSA_Tradeoff}, we see that the overclocked RCA achieves the fastest operating frequencies under most area constraints when the accuracy requirement is released.

@@

\subsection{Experimental Setup} \label{sub:experimental_setup}
We initially build up a test framework on an FPGA. The general architecture is depicted in Fig.~\ref{Test Framework}. The main body of the test framework consists of the circuit under test (CUT), the test frequency generator and the control logic, as shown in the dotted box in Fig.~\ref{Test Framework}. The I/Os of the CUT are registered by the launch registers (LRs) and the sample registers (SRs), which are all triggered by the test clock. Input test vectors are stored in the on-chip memory during initialization. The results are sampled using Xilinx ChipScope. Finally, we perform an offline comparison of the output of the original circuit at the rated frequency with the output of the overclocked as well as the truncated designs using the same input vectors.
%
\begin{figure}[htbp]
  \centering
  \vspace{-5ex}
  \includegraphics[width=3.5in]{./Figures/TestFramework.pdf}
  \vspace{-3ex}
  \caption{Test framework, which is composed of a measurement architecture (the dotted box) on an FPGA and an off-line comparator using software. The error-free data are obtained by either pre-computation or initial run with low frequencies.}
  \label{Test Framework}
  \vspace{-1.5ex}
\end{figure}

@@

For example, the predicted frequencies and the actual frequencies of a $5^{th}$~order FIR filter using different word-lengths are depicted in Fig.~\ref{Fig:TimingMargin}. The ``actual'' maximum frequencies are computed by increasing the operating frequency from the rated value until errors are observed at the output; the maximum operating frequency with correct output is recorded for the current word-length. As can be seen in Fig.~\ref{Fig:TimingMargin}, the circuit can operate without errors at a much higher frequency in practice than predicted according to our experiments. A maximum speed differential of $3.2\times$ is obtained when the input signal is 5-bit.
%
\begin{figure}[ht]
  \centering
  \vspace{-2ex}
  \includegraphics[width=3.2in]{./Figures/Exp/FIR/RatedFrequency3.eps}
  \vspace{-1ex}
  \caption{The maximum operating frequencies for different input word-lengths of an FIR filter. The dotted line depicts the rated frequency reported by the timing analysis tool. The solid line is obtained through real FPGA tests using our platform.}
  \label{Fig:TimingMargin}
  \vspace{-.5ex}
\end{figure}

@@

\section{Results and Discussion}\label{Section_Experiments}
\subsection{Case study: FIR filter}
We first assess the accuracy of our proposed models of errors. We record the value of MRE based on (\ref{Eq:MRE}) in a $5^{th}$~order FIR with respect to different operating frequencies. The modeled values of both overclocking error and truncation error of the FIR filter are presented in Fig.~\ref{Fig:FIR} (dotted lines), as well as the actual measurements on the FPGA (solid lines) with two types of input data. The results demonstrate that our models match well with the practical results obtained using the uniform independent inputs.
% FIR
\begin{figure}[tbp]
  \centering
  \vspace{-3ex}
  \includegraphics[width=3.5in]{./Figures/Exp/FIR/FIR_Error_New_TCAD.eps}
  \vspace{-2ex}
  \caption{A demonstration of two design perspectives with a $5^{th}$~order FIR filter, which is implemented on Virtex-6 FPGA. The modeled values of both overclocking errors and truncation errors are presented as dotted lines. The actual FPGA measurements are depicted using solid lines. Two types of inputs are employed in the overclocking scenario: the uniformly distributed data and the real image data from Lena.}
  \label{Fig:FIR}
\end{figure}

\begin{figure}[ht]
  \vspace{-2ex}
  % 425M, 450M, 480M, 520M
  \centering
  \subfigure[\scriptsize n=8,~SNR=47.97dB.]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_T8.eps}
  }% This is important! use % to indicate same line, otherwise new line
  \subfigure[\scriptsize n=8,~SNR=44.94dB.]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_T65.eps}
  }%
  \subfigure[\scriptsize n=8,~SNR=21.58dB]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_T62.eps}
  }%
  \subfigure[\scriptsize n=8,~SNR=10.03dB]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_T6.eps}
  }
  \subfigure[\scriptsize n=7,~SNR=26.02dB]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_Q7T10.eps}
  }% This is important! use % to indicate same line, otherwise new line
  \subfigure[\scriptsize n=5,~SNR=21.76dB]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_Q4T10.eps}
  }%
  \subfigure[\scriptsize n=2,~SNR=6.57dB]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_Q2T10.eps}
  }%
  \subfigure[\scriptsize n=1,~SNR=3.95dB]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_Q1T10.eps}
  }
\vspace{-2ex}
\caption{Output images of the FIR filter for both overclocking scenario (top row) and traditional scenario (bottom row) when operating frequencies are $425MHz$, $450MHz$, $480MHz$ and $520MHz$ respectively (from left to right).}
\label{Fig:FIR_Lena}
\vspace{-1.5ex}
\end{figure}

According to Fig.~\ref{Fig:FIR}, output errors are reduced in the overclocking scenario for both input types in comparison to the traditional scenario, as the analytical model validates. In addition, we see that using real data, more significant reduction of MRE are achieved, and that no errors are observed when frequency is initially increased. This is because for real data, long carry chains are typically generated with even smaller probabilities, and the longest carry chain rarely occurs.

@@

\subsection{Potential Benefits in Datapath Design}
As we mentioned in Section~\ref{Section_OptimumAdder}, our results could be of interest to a circuit designer in two ways. Typically, either the designer will want to create a circuit that can run at a given frequency with the minimum possible MRE, or the algorithm designer will wish to run as fast as possible whilst maintaining a specific error tolerance. For the first design target, the experimental results for all five example designs on FPGA are summarized in Table~\ref{Table:Ratio_MRE} in terms of the relative reduction of MRE as given in~(\ref{Eq:Relative_Reduction_MRE}) where $MRE_{Trad}$ and $MRE_{ovrc}$ denote the value obtained in the traditional scenario and in the overclocking scenario, respectively.
%
\begin{eqnarray}\label{Eq:Relative_Reduction_MRE}
  \frac{MRE_{Trad}-MRE_{ovrc}}{MRE_{Trad}}\times 100\%
\end{eqnarray}

In this table, the frequency is normalized to the maximum error-free frequency for each design when the input signal is 8-bit. The N/A in Table~\ref{Table:Ratio_MRE} refers to the situations where a certain frequency simply cannot be achieved using the traditional scenario. It can be seen that a significant reduction of MRE can be achieved using the proposed overclocking scenario, and the geometric mean reduction varies from $67.9\%$ to $95.4\%$ using uniform input data. Even larger differences of MRE can be observed when testing with real image data for each design, ranging from $83.6\%$ to $98.8\%$, as expected given the results shown in Fig.~\ref{Fig:FIR}.
%
\begin{table*}[tbp]
\vspace{-1ex}
 %% increase table row spacing, adjust to taste
   \renewcommand{\arraystretch}{1.05}
   \setlength{\tabcolsep}{3.5pt}
   \tbl{Relative Reduction of MRE ($\%$) in Overclocking Scenario for Various Normalized Frequencies Based on~(\ref{Eq:Relative_Reduction_MRE}) for Two Types of Input Data: the Uniform Data (Uni) and Real Image Data from Lena (Lena).}{
   \centering
   %\footnotesize
   \scriptsize
   %\vspace{-4mm}
   \begin{threeparttable}
   \begin{tabular}{|c|cc|cc|cc|cc|cc||cc|}
     \hline
     \multirow{2}*{\begin{tabular}{c}\textbf{Normalized}\\\textbf{Frequency}\end{tabular}} &
     \multicolumn{2}{c|}{\textbf{FIR}} &
     \multicolumn{2}{c|}{\textbf{Sobel}} &
     \multicolumn{2}{c|}{\textbf{IIR}} &
     \multicolumn{2}{c|}{\textbf{Butterworth}} &
     \multicolumn{2}{c||}{\textbf{DCT4}} &
     \multicolumn{2}{c|}{\textbf{Geo.Mean}}\\
     &  {Uni} & {Lena } &  {Uni} & {Lena } &  {Uni} & {Lena }
     &  {Uni} & {Lena } &  {Uni} & {Lena } &  {Uni} & {Lena }\\
     \hline
     1.04 & 99.85	& 100.00  & 99.51 &	99.74 &	 72.28 &	90.09 &	 79.03 &	 100.00 &	 83.58 &	 98.06 & 86.14 & 97.50\\
     1.08 & 98.93	& 99.97   & 96.26 &	93.75 &	 71.64 &	90.50 &	 78.81 &	 100.00 &	 83.26 &	 98.45 & 85.15 & 96.46\\
     1.12 & 94.27	& 98.82   & 96.25 &	93.62 &	 73.63 &   88.25 &	 81.88 &	 84.87  &	 89.44 &	 99.56 & 86.68 & 92.84\\
     1.16 & 99.66	& 99.91   & 73.73 &	93.62 &	 73.10 &   89.92 &	 79.30 &	 84.23  &	 79.07 &	 99.44 & 80.44 & 93.23\\
     1.20 & 96.03	& 99.90   & 81.55 &	81.52 &	 70.76 &	75.67 &	 64.96 &	 84.50  &	 N/A\tnote{*}	  & N/A\tnote{*} & 77.46 & 84.95 \\
     1.24 & 98.46	& 99.32   & 81.43 &	81.67 &	 70.47 &	76.12 &	 63.66 &	 84.23  &	 N/A\tnote{*}	  & N/A\tnote{*} & 77.44 & 84.92\\
     1.28 & 95.39	& 99.29   & 60.41 &	78.24 &	 N/A\tnote{*}	    &   N/A\tnote{*}     &	54.38 &	 75.15  &	N/A\tnote{*}	  & N/A\tnote{*} & 67.92 & 83.58\\
     1.32 & 95.37	& 98.75   & N/A\tnote{*}	  &	 N/A\tnote{*}	    &	 N/A\tnote{*}	    &	 N/A\tnote{*}   & N/A\tnote{*}   & N/A\tnote{*}    &    N/A\tnote{*}   & N/A\tnote{*} & 95.37	& 98.75 \\
     \hline
   \end{tabular}
   \normalsize
   \begin{tablenotes}
   \scriptsize
    \item[*] Current frequency cannot be achieved in the traditional scenario. These points are excluded from the calculation of geometric means.
   \end{tablenotes}
   \end{threeparttable}}
   \label{Table:Ratio_MRE}
 %\vspace{-1ex}
 \end{table*}

>>>>

\keywords{Overclocking, Imprecise Design, Approximate Computing, Truncation}

@@

\section{Background}\label{Section_Background}
\subsection{Word-length Optimization}\label{Section_Background_WL}
There exists a significant amount of work demonstrating that optimizing the datapath precision would bring a substantial benefits on clock speed, silicon area and power consumption. There are two main research directions in this area. The first discusses methods to analyze the errors caused by the usage of finite precision and compute bounds on the worst-case errors seen at the output of the design. Related techniques include the simulation-based approach~\cite{SungWLop_Simulation} and the analytical-based approach such as interval arithmetic~\cite{moore1966IA}, affine arithmetic~\cite{de2004affine}, Satisfiability-Modulo Theories~\cite{SAT2010} and polynomial representations~\cite{Boland2011TCAD}. A detailed summary of these methods can be found in~\cite{GeorgeDT11}. In the second research field, these tools are utilized to choose the minimum precision within a datapath to satisfy a required accuracy specification~\cite{RoldaoWLop_Simulation}. In this work, two different design scenarios are analyzed. In the conventional scenario, we speculate and reduce the original word-length in the datapath under a given timing constraint while ensuring no timing errors will occur. In the proposed new scenario, we operate the datapath with the original word-length while allowing timing violations sometimes to happen. The truncation errors derived in the former scenario are then compared to the overclocking errors generated in our proposed new scenario.

@@

\subsection{Imprecise Circuit Design Methodology}\label{Section_Background_Imprecise}
It is worth noting that the choice of precision is not the only source of error when designing a datapath. Recently we have seen a growth of parallel streams of research which aim at exploring alternative methods to trade accuracy for design efficiency. This strand of research is motivated by the fact that extra benefits of manufacturing, test, power and timing can be obtained if the tight requirement of absolute correctness is released for devices and interconnects, as pointed out by the International Technology Roadmap for Semiconductors (ITRS) in 2007~\cite{semiconductor2007international}. This topic is expected to be of growing importance in the future, because we will face new design challenges as the technology scales further.

@@

\subsection{Determination of the Optimum Adder Structure for Given Frequency Requirements}
First, suppose the algorithm designer wished to create a circuit that can run at a given frequency in a given resource constraint while achieving the minimum possible output error. A decision must then be taken over which adder structure achieves this minimum, and whether the structure should be overclocked.

@@

We first notice from Fig.~\ref{Fig_CSA_LUT25} that for all frequency values, the overclocked RCA achieves no larger error expectation than the RCA with truncated operand word-lengths, as predicted by the models for the RCA in Section~\ref{section_RCA_Comparison}. It can also be observed that the CSA cannot be implemented with the original word-length due to the limited area budget and this leads to large truncation errors. Although the CSA with 4 stages is best for some frequencies, the overclocked RCA is still the optimum design when high operating frequencies are applied.

We then perform similar experiments with a variety of area constraints. The optimum design methods with respect to different operating frequencies and area consumptions are demonstrated in Fig.~\ref{Fig_CSA_Tradeoff}. From this figure, several observations can be made. Firstly, if the frequency requirement is not large whilst the available area is large enough to implement a CSA in full precision, it will be the optimum design. This is expected from our earlier analysis in Fig.~\ref{Fig_CSA2stage_Timing}. Secondly, the 2-stage CSA is better than the 3-stage CSA and the 4-stage CSA when frequency is low, as it consumes less area although all achieve the same error expectation. Thirdly, it is possible that only part of the CSA can be implemented under a tighter area budget, whereas the RCA still keeps full precision. In this case, area becomes the dominant factor and precision is lost for the CSA, meaning the RCA with overclocking is the optimum design method across almost the whole frequency domain in this situation. Finally, for very stringent area constraints, the word-length of RCA is also limited. This results in truncation errors initially for all design scenarios. However, the RCA with overclocking can still be employed as the optimum design, because it loses less precision than the CSA under the same area constraint.
%
\begin{figure}[t]
  \centering
  \vspace{-2.5ex}
  \includegraphics[width=4.5in]{./Figures/Tradeoff_Freq.eps}
  \vspace{-2.5ex}
  \caption{A demonstration of the optimum design methodology which achieves the minimum error at outputs with respect to a variety of frequency and area constraints.}
  \label{Fig_CSA_Tradeoff}
\end{figure}

\subsection{Determination of the Optimum Adder Structure for Given Accuracy Requirements}
If the design goal is to operate the circuit as fast as possible with the minimum area whilst a certain error budget can be tolerated, the optimum design methodology can be decided as illustrated in Fig.~\ref{Fig_CSA_Tradeoff_Error}. In this situation, the error specifications are evaluated in terms of the mean relative error ($MRE$), as given by~(\ref{Eq:MRE}), where $E_{error}$ and $E_{out}$ refer to the mean value of error and the mean value of outputs, respectively.
%
\begin{eqnarray}\label{Eq:MRE}
  MRE=\left|\frac{E_{error}}{E_{out}}\right|\times 100\%
\end{eqnarray}

In our experiments, $MRE$ is set ranging from $0.001\%$ to $50\%$. For a certain value of $MRE$, the design with the maximum operating frequency is selected as the optimum design. Moreover, the smallest design is the optimum one if multiple structures operate at the same frequency, with a certain area requirement. Fig.~\ref{Fig_CSA_Tradeoff_Error} can thus be obtained based on these criteria.
%
\begin{figure}[t]
    \centering
    \vspace{-2ex}
    \includegraphics[width=4.5in]{./Figures/Tradeoff_Error2.eps}
    \vspace{-2.5ex}
    \caption{A demonstration of the optimum design methodology which runs at the fastest frequency with respect to a variety of accuracy and area constraints.}
    \label{Fig_CSA_Tradeoff_Error}
    \vspace{-1ex}
\end{figure}

For a tight accuracy requirement, e.g. $MRE<0.005\%$, CSA serves as the optimum design choice for large area constraints, as it intrinsically operates faster than RCA. Once again, when the area budget shrinks, the RCA performs best because the precision of the CSA is limited. Similarly to the results in Fig.~\ref{Fig_CSA_Tradeoff}, we see that the overclocked RCA achieves the fastest operating frequencies under most area constraints when the accuracy requirement is released.

@@

\subsection{Experimental Setup} \label{sub:experimental_setup}
We initially build up a test framework on an FPGA. The general architecture is depicted in Fig.~\ref{Test Framework}. The main body of the test framework consists of the circuit under test (CUT), the test frequency generator and the control logic, as shown in the top dotted box in Fig.~\ref{Test Framework}. The I/Os of the CUT are registered by the launch registers (LRs) and the sample registers (SRs), which are all triggered by the test clock. Input test vectors are stored in the on-chip memory during initialization. The results are sampled using Xilinx ChipScope. Finally, we perform an offline comparison of the output of the original circuit at the rated frequency with the output of the overclocked as well as the truncated designs using the same input vectors.
%
\begin{figure}[htbp]
  \centering
  \vspace{-3.5ex}
  \includegraphics[width=3.7in]{./Figures/TestFramework.pdf}
  \vspace{-1.5ex}
  \caption{Test framework, which is composed of a measurement architecture (the top dotted box) on an FPGA and an off-line comparator using software. The error-free data are obtained by either pre-computation or initial run with low frequencies.}
  \label{Test Framework}
  \vspace{-1.5ex}
\end{figure}

@@

For example, the frequencies predicted by Xilinx ISE and the actual frequencies of a $5^{th}$~order FIR filter using different word-lengths are depicted in Fig.~\ref{Fig:TimingMargin}. The ``actual'' maximum frequencies are computed by increasing the operating frequency from the rated value until errors are observed at the output; the maximum operating frequency with correct output is recorded for the current word-length. As can be seen in Fig.~\ref{Fig:TimingMargin}, the circuit can operate without errors at a much higher frequency in practice than predicted according to our experiments. A maximum speed differential of $3.2\times$ is obtained when the input signal is 5-bit.
%
\begin{figure}[ht]
  \centering
  \vspace{-2ex}
  \includegraphics[width=3.2in]{./Figures/Exp/FIR/RatedFrequency3.eps}
  \vspace{-1ex}
  \caption{The maximum operating frequencies for different input word-lengths of an FIR filter. The dotted line depicts the rated frequency reported by the timing analysis tool. The solid line is obtained through real FPGA tests using our platform.}
  \label{Fig:TimingMargin}
  \vspace{-.5ex}
\end{figure}

@@

\section{Results and Discussion}\label{Section_Experiments}
\subsection{Case study: FIR filter}
We first assess the accuracy of our proposed models of errors. We record the value of $MRE$ based on (\ref{Eq:MRE}) in a $5^{th}$~order FIR with respect to different operating frequencies. The modeled values of both overclocking error and truncation error of the FIR filter are presented in Fig.~\ref{Fig:FIR} (dotted lines), as well as the actual measurements on the FPGA (solid lines) with two types of input data. The results demonstrate that our models match well with the practical results obtained using the uniform independent inputs.
% FIR
\begin{figure}[tbp]
  \centering
  \vspace{-3ex}
  \includegraphics[width=3.5in]{./Figures/Exp/FIR/FIR_Error_New_TCAD.eps}
  \vspace{-2ex}
  \caption{A demonstration of two design perspectives with a $5^{th}$~order FIR filter, which is implemented on Virtex-6 FPGA. The modeled values of both overclocking errors and truncation errors are presented as dotted lines. The actual FPGA measurements are depicted using solid lines. Two types of inputs are employed in the overclocking scenario: the uniformly distributed data and the real image data from Lena.}
  \label{Fig:FIR}
\end{figure}

\begin{figure}[ht]
  \vspace{-2ex}
  % 425M, 450M, 480M, 520M
  \centering
  \subfigure[\scriptsize n=8,~SNR=47.97dB.]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_T8.eps}
  }% This is important! use % to indicate same line, otherwise new line
  \subfigure[\scriptsize n=8,~SNR=44.94dB.]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_T65.eps}
  }%
  \subfigure[\scriptsize n=8,~SNR=21.58dB.]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_T62.eps}
  }%
  \subfigure[\scriptsize n=8,~SNR=10.03dB.]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_T6.eps}
  }
  \subfigure[\scriptsize n=7,~SNR=26.02dB.]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_Q7T10.eps}
  }% This is important! use % to indicate same line, otherwise new line
  \subfigure[\scriptsize n=5,~SNR=21.76dB.]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_Q4T10.eps}
  }%
  \subfigure[\scriptsize n=2,~SNR=6.57dB.]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_Q2T10.eps}
  }%
  \subfigure[\scriptsize n=1,~SNR=3.95dB.]{
    \includegraphics[width=.24\textwidth]{./Figures/Exp/FIR/LenaImages/FIRLena_Q1T10.eps}
  }
\vspace{-2ex}
\caption{Output images of the FIR filter for both overclocking scenario (top row) and traditional scenario (bottom row) when operating frequencies are $425MHz$, $450MHz$, $480MHz$ and $520MHz$ respectively (from left to right).}
\label{Fig:FIR_Lena}
\vspace{-1.5ex}
\end{figure}

According to Fig.~\ref{Fig:FIR}, output errors are reduced in the overclocking scenario for both input types in comparison to the traditional scenario, as the analytical model validates. In addition, we see that using real data, more significant reduction of $MRE$ are achieved, and that no errors are observed when frequency is initially increased. This is because for real data, long carry chains are typically generated with even smaller probabilities, and the longest carry chain rarely occurs.

@@

\subsection{Potential Benefits in Datapath Design}
As we mentioned in Section~\ref{Section_OptimumAdder}, our results could be of interest to a circuit designer in two ways. Typically, either the designer will want to create a circuit that can run at a given frequency with the minimum possible $MRE$, or the algorithm designer will wish to run as fast as possible whilst maintaining a specific error tolerance. For the first design target, the experimental results for all five example designs on FPGA are summarized in Table~\ref{Table:Ratio_MRE} in terms of the relative reduction of $MRE$ as given in~(\ref{Eq:Relative_Reduction_MRE}) where $MRE_{Trad}$ and $MRE_{ovrc}$ denote the value obtained in the traditional scenario and in the overclocking scenario, respectively.
%
\begin{eqnarray}\label{Eq:Relative_Reduction_MRE}
  \frac{MRE_{Trad}-MRE_{ovrc}}{MRE_{Trad}}\times 100\%
\end{eqnarray}

In this table, the frequency is normalized to the maximum error-free frequency for each design when the input signal is 8-bit. The N/A in Table~\ref{Table:Ratio_MRE} refers to the situations where a certain frequency simply cannot be achieved using the traditional scenario. It can be seen that a significant reduction of $MRE$ can be achieved using the proposed overclocking scenario, and the geometric mean reduction varies from $67.9\%$ to $95.4\%$ using uniform input data. Even larger differences of $MRE$ can be observed when testing with real image data for each design, ranging from $83.6\%$ to $98.8\%$, as expected given the results shown in Fig.~\ref{Fig:FIR}.
%
\begin{table*}[tbp]
\vspace{-1ex}
 %% increase table row spacing, adjust to taste
   \renewcommand{\arraystretch}{1.05}
   \setlength{\tabcolsep}{3.5pt}
   \tbl{Relative Reduction of MRE ($\%$) in Overclocking Scenario for Various Normalized Frequencies Based on~(\ref{Eq:Relative_Reduction_MRE}) for Two Types of Input Data: Uniform Data (Uni) and Real Image Data from Lena (Lena).}{
   \centering
   %\footnotesize
   \scriptsize
   %\vspace{-4mm}
   \begin{threeparttable}
   \begin{tabular}{|c|cc|cc|cc|cc|cc||cc|}
     \hline
     \multirow{2}*{\begin{tabular}{c}\textbf{Normalized}\\\textbf{Frequency}\end{tabular}} &
     \multicolumn{2}{c|}{\textbf{FIR}} &
     \multicolumn{2}{c|}{\textbf{Sobel}} &
     \multicolumn{2}{c|}{\textbf{IIR}} &
     \multicolumn{2}{c|}{\textbf{Butterworth}} &
     \multicolumn{2}{c||}{\textbf{DCT4}} &
     \multicolumn{2}{c|}{\textbf{Geo.Mean}}\\
     &  {Uni} & {Lena } &  {Uni} & {Lena } &  {Uni} & {Lena }
     &  {Uni} & {Lena } &  {Uni} & {Lena } &  {Uni} & {Lena }\\
     \hline
     1.04 & 99.85	& 100.00  & 99.51 &	99.74 &	 72.28 &	90.09 &	 79.03 &	 100.00 &	 83.58 &	 98.06 & 86.14 & 97.50\\
     1.08 & 98.93	& 99.97   & 96.26 &	93.75 &	 71.64 &	90.50 &	 78.81 &	 100.00 &	 83.26 &	 98.45 & 85.15 & 96.46\\
     1.12 & 94.27	& 98.82   & 96.25 &	93.62 &	 73.63 &   88.25 &	 81.88 &	 84.87  &	 89.44 &	 99.56 & 86.68 & 92.84\\
     1.16 & 99.66	& 99.91   & 73.73 &	93.62 &	 73.10 &   89.92 &	 79.30 &	 84.23  &	 79.07 &	 99.44 & 80.44 & 93.23\\
     1.20 & 96.03	& 99.90   & 81.55 &	81.52 &	 70.76 &	75.67 &	 64.96 &	 84.50  &	 N/A\tnote{*}	  & N/A\tnote{*} & 77.46 & 84.95 \\
     1.24 & 98.46	& 99.32   & 81.43 &	81.67 &	 70.47 &	76.12 &	 63.66 &	 84.23  &	 N/A\tnote{*}	  & N/A\tnote{*} & 77.44 & 84.92\\
     1.28 & 95.39	& 99.29   & 60.41 &	78.24 &	 N/A\tnote{*}	    &   N/A\tnote{*}     &	54.38 &	 75.15  &	N/A\tnote{*}	  & N/A\tnote{*} & 67.92 & 83.58\\
     1.32 & 95.37	& 98.75   & N/A\tnote{*}	  &	 N/A\tnote{*}	    &	 N/A\tnote{*}	    &	 N/A\tnote{*}   & N/A\tnote{*}   & N/A\tnote{*}    &    N/A\tnote{*}   & N/A\tnote{*} & 95.37	& 98.75 \\
     \hline
   \end{tabular}
   \normalsize
   \begin{tablenotes}
   \scriptsize
    \item[*] Current frequency cannot be achieved in the traditional scenario. These points are excluded from the calculation of geometric means.
   \end{tablenotes}
   \end{threeparttable}}
   \label{Table:Ratio_MRE}
 %\vspace{-1ex}
 \end{table*}
